---
title: "Hotels"
author: "Kristen Sohm"
date: "4/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(rpart)
library(verification)
library(klaR)
library(dplyr)
```

```{r}
kappa=function(x){
  n=sum(x)
  pobs=(x[1,1]+x[2,2])/n
  pexp=(sum(x[1,])*sum(x[,1])+sum(x[2,])*sum(x[,2]))/n^2
  kappa=(pobs-pexp)/(1-pexp)
  t1=0
  t2=0
  t3=0
  pii=x/n
  pidot=apply(pii,1,sum)
  pdotj=apply(pii,2,sum)
  for(i in 1:2){
    t1 = t1 + pii[i,i]*((1-pexp) - (1-pobs)*(pidot[i]+pdotj[i]))^2
  }
  t2 = pii[1,2]*(pdotj[1]+pidot[2])^2 + pii[2,1]*(pdotj[2] + pidot[1])^2
  t3 = (pobs*pexp-2*pexp+pobs)^2
  vhat = (t1 + t2*(1-pobs)^2 -t3)/(n*(1-pexp)^4)
  se=sqrt(vhat)
  return(c(kappa,se))
}


class.sum=function(truth,predicted){
  xt=table(truth,round(predicted+0.000001))
  pcc=round(100*sum(diag(xt))/sum(xt),2)
  spec=round(100*xt[1,1]/sum(xt[1,]),2)
  sens=round(100*xt[2,2]/sum(xt[2,]),2)
  kap=round(kappa(xt)[1],4)
  au=round(roc.area(truth,predicted)$A,4)
  return(cbind(c("Percent Correctly Classified = ","Specificity = ","Sensitivity = ","Kappa =","AUC= "),c(pcc,spec,sens,kap,au)))
}

```
Variables for Hotel Data:
```{r}
HotelData=read.csv('C:/Users/krist/Documents/STAT5650/data/hotel_bookings.csv')
#HotelData
colnames(HotelData)
```

```{r}
Hotel = names(HotelData) %in% c("hotel")
HotelDataSansHotel= HotelData[!Hotel]
```
Distributions

Normal Q-Q Plots
```{r}
par(mar=c(.5,.5,.5,.5))
par(mfrow=c(13,2))
qqnorm(HotelDataSansHotel$lead_time, main="lead_time")
qqnorm(HotelDataSansHotel$arrival_date_year, main="arrival_date_year")
#qqnorm(HotelDataSansHotel$arrival_date_month, main="lead_time")
qqnorm(HotelDataSansHotel$arrival_date_week_number, main="arrival_date_week_number")
qqnorm(HotelDataSansHotel$arrival_date_day_of_month, main="arrival_date_day_of_month")

qqnorm(HotelDataSansHotel$stays_in_weekend_nights, main="stays_in_weekend_nights")
qqnorm(HotelDataSansHotel$stays_in_week_nights, main="stays_in_week_nights")
qqnorm(HotelDataSansHotel$adults, main="adults")
qqnorm(HotelDataSansHotel$children, main="children")
qqnorm(HotelDataSansHotel$babies, main="babies")

#qqnorm(HotelDataSansHotel$meal)
#qqnorm(HotelDataSansHotel$market_segment)
#qqnorm(HotelDataSansHotel$distribution_channel)
qqnorm(HotelDataSansHotel$is_repeated_guest, main="is_repeated_guest")
qqnorm(HotelDataSansHotel$previous_cancellations, main="previous_cancellations")

qqnorm(HotelDataSansHotel$previous_bookings_not_canceled, main="previous_bookings_not_canceled")
#qqnorm(HotelDataSansHotel$reserved_room_type)
#qqnorm(HotelDataSansHotel$assigned_room_type)
qqnorm(HotelDataSansHotel$booking_changes, main="booking_changes")
#qqnorm(HotelDataSansHotel$deposit_type)

qqnorm(HotelDataSansHotel$days_in_waiting_list, main="days_in_waiting_list")
#qqnorm(HotelDataSansHotel$customer_type)
qqnorm(HotelDataSansHotel$adr, main="adr")
qqnorm(HotelDataSansHotel$required_car_parking_spaces, main="required_car_parking_spacese")
qqnorm(HotelDataSansHotel$total_of_special_requests, main="total_of_special_requests")

```
Histograms
```{r}
library(rcompanion)
par(mar=c(1,1,1,1))
par(mfrow=c(9,3))

plotNormalHistogram(HotelDataSansHotel$lead_time, main="lead_time")
plotNormalHistogram(HotelDataSansHotel$arrival_date_year, main="arrival_date_year")
plotNormalHistogram(HotelDataSansHotel$arrival_date_week_number, main="arrival_date_week_number")

plotNormalHistogram(HotelDataSansHotel$arrival_date_day_of_month, main="arrival_date_day_of_month")
plotNormalHistogram(HotelDataSansHotel$stays_in_weekend_nights, main="stays_in_weekend_nights")
plotNormalHistogram(HotelDataSansHotel$stays_in_week_nights, main="stays_in_week_nights")

plotNormalHistogram(HotelDataSansHotel$adults, main="adults")
plotNormalHistogram(HotelDataSansHotel$children, main="children")
plotNormalHistogram(HotelDataSansHotel$babies, main="babies")

plotNormalHistogram(HotelDataSansHotel$is_repeated_guest, main="is_repeated_guest")
plotNormalHistogram(HotelDataSansHotel$previous_cancellations, main="previous_cancellations")
plotNormalHistogram(HotelDataSansHotel$previous_bookings_not_canceled, main="previous_bookings_not_canceled")

plotNormalHistogram(HotelDataSansHotel$booking_changes, main="booking_changes")
plotNormalHistogram(HotelDataSansHotel$days_in_waiting_list, main="days_in_waiting_list")
plotNormalHistogram(HotelDataSansHotel$adr, main="adr")

plotNormalHistogram(HotelDataSansHotel$required_car_parking_spaces, main="required_car_parking_spacese")
plotNormalHistogram(HotelDataSansHotel$total_of_special_requests, main="total_of_special_requests")


```

Variables with binomial or numeric values:
```{r}
#subset to include only binomial and numeric variables
nonNumericVariables = names(HotelData) %in% c("hotel", "arrival_date_month", "meal", "country", "market_segment", "distribution_channel", "reserved_room_type", "assigned_room_type", "deposit_type", "agent", "company", "customer_type", "reservation_status", "reservation_status_date")
HotelDataSubset=HotelData[!nonNumericVariables]
colnames(HotelDataSubset)

```


```{r}
#drop rows with missing values (4 NAN values in children attribute)
library(tidyr)
HotelDataNumericSansNAN = drop_na(HotelDataSubset)
```

**Predict hotel reservation cancelations**
Logistic Regression: numeric and binomial predictor variables, is_canceled as response variable.
```{r}
#logistic regression model
HotelDataNumeric.lr = glm(is_canceled~ . ,family=binomial,data=HotelDataNumericSansNAN)

#confusion matrix and accuracy (Dr. Cutler's way)
table(HotelDataNumericSansNAN$is_canceled,round(predict(HotelDataNumeric.lr,type="response")))
class.sum(HotelDataNumericSansNAN$is_canceled,predict(HotelDataNumeric.lr,type="response"))
```
```{r}
#library(caret)
##glm model
#HotelDataNumeric.lr = glm(is_canceled~ . , data=HotelDataNumericSansNAN,family=binomial())
##use model to make predictions
#glub = predict(HotelDataNumeric.lr, type ="response")
##display confusion matrix and accuracy
#confusionMatrix(data = glub, reference = HotelDataNumericSansNAN$is_canceled)
```

Stepwise variable selection:
```{r}
#logistic regression model
HotelDataNumeric.lr = glm(is_canceled~ . ,family=binomial,data=HotelDataNumericSansNAN)
step.model=stepAIC(HotelDataNumeric.lr, direction = "both", trace = FALSE)
summary(step.model)
step.model
class.sum(HotelDataNumericSansNAN$is_canceled,predict(step.model,type="response"))
```
Logistic Regression with Variables Selected with Stepwise AIC.
```{r}
#logistic regression model
HotelDataNumeric.lr = glm(is_canceled~ lead_time + arrival_date_year + arrival_date_week_number + 
    stays_in_weekend_nights + adults + children + babies + is_repeated_guest + 
    previous_cancellations + previous_bookings_not_canceled + 
    booking_changes + days_in_waiting_list + adr + required_car_parking_spaces + 
    total_of_special_requests ,family=binomial,data=HotelDataNumericSansNAN)

#confusion matrix and accuracy (Dr. Cutler's way)
table(HotelDataNumericSansNAN$is_canceled,round(predict(HotelDataNumeric.lr,type="response")))
class.sum(HotelDataNumericSansNAN$is_canceled,predict(HotelDataNumeric.lr,type="response"))
```

Cross-validated Logistic Regression with Variables from Stepwise AIC:
```{r}
HotelDataNumeric.lr.xval=rep(0,nrow(HotelDataNumericSansNAN))
xvs=rep(1:10,length=nrow(HotelDataNumericSansNAN))
xvs=sample(xvs)
for(i in 1:10){
    train=HotelDataNumericSansNAN[xvs!=i,]
    test=HotelDataNumericSansNAN[xvs==i,]
    glub=glm(is_canceled~ lead_time + arrival_date_year + arrival_date_week_number + 
    stays_in_weekend_nights + adults + children + babies + is_repeated_guest + 
    previous_cancellations + previous_bookings_not_canceled + 
    booking_changes + days_in_waiting_list + adr + required_car_parking_spaces + 
    total_of_special_requests ,family=binomial,data=train)
    HotelDataNumeric.lr.xval[xvs==i]=predict(glub,test,type="response")
}
table(HotelDataNumericSansNAN$is_canceled,round(HotelDataNumericSansNAN.lr.xval))
class.sum(HotelDataNumericSansNAN$is_canceled,HotelDataNumericSansNAN.lr.xval)
```

***Are resort hotels and city hotels significantly different enough to warrant different prediction models?***
```{r}
#separate resort hotel rows from city hotel rows
resortHotel=HotelData[1:40060,1:32]
cityHotel=HotelData[40061:119390,1:32]
summary(resortHotel)
```

```{r}
unique(resortHotel$hotel)
```


```{r}
summary(cityHotel)
```

